### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/temporal.tex
Model Section moved to temporal file

In this sense the action of posting is when
users want to put something on their wall who want others to know through
their social networks. Likewise the storing action occurs when users want to
retain permanently information that are of their interest. Finally the
browsing action occurs when users are exploring content for their needs. All
these actions go beyond only evaluating individuals.

\begin{figure*}
  \captionsetup{justification=centering,margin=2cm}
  \centering
  \setlength\fboxsep{0pt}
  \setlength\fboxrule{0.7pt}
  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/users.png}}
  \caption{Users actions on interactive evolutionary systems.}
  \label{fig:users}       
\end{figure*}

For the evaluating action is proposed  that users evaluate individuals
indirectly, this means users can evaluate accepting indirect recommendations of
friends that are currently active in the system. These recommendations can be
store individuals in the system of users who know each other.

Also the  browsing action is proposed within IEC systems,  this means that users
can explore information that other active users in the system are generating.

Finally activity participation is proposed. This can be divided into four
different actions as fallows:

\begin{itemize} 
\item The Inviting action occurs when a user invites another to the system through social networks or from person to person.  
\item The sharing action occurs when users share their individual creations with others active user in the system.
\item The posting action occurs when published what they are doing within the system.
\item The storing action occurs when users keep individuals in a collection, the collections concept will be discussed later in this chapter.
\end{itemize}

All the mention above is shown in figure \ref{fig:users}

\subsection{Individual}

Individuals are entities that form the population for the evolutionary
algorithm. In particular for  this work individuals are animated digital
paintings, every individual in the population is defined by a chromosome [].
This chromosome defines the behavior of the individual, so that it consists of a
vector of real numbers of fifteen genes, where each gen define a particular
behavior in painting. This individuals combined each other to generated new
individual in the population following the genetic operators such as selection,
crossover and mutation[].

\subsection{User Model} Now that both users an individuals have been defined we can
model the users behavior have with respect to individuals in a graph-based user
model.

The reason for using a graph to model the user-individual is because it could
can express in a simple way the behavior of this two entities having the
necessary information  to represent the knowledge,  for example a vertex user
and individual vertex connected through a edge that  represents the
relationship. this relationship can be the semantic  between these two entities
as figure \ref{fig:User-Individual} show. Each of these vertices contain properties as well as the
edges, these properties will be explained as fallows.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_individual.png}}
\caption{User-Individual.}
\label{fig:User-Individual}       
\end{figure*}

In this proposed graph-based user-individual model the vertices are represented
by set of nodes of USERS, INDIVIDUALS and COLLECTIONS. The edges are define by a
set of relationships as LIKES, KNOWS, PARENT, HAS that represents the
relationships between the vertices as seen in figure \ref{fig:User-Individual}.

Already explained that represent USERS and INDIVIDUALS, now an explanation is
given of what concerns for the concept of COLLECTIONS.

In this proposal a collection is defined as a  container where users can store
selected individuals laid within the collection. A collection is created when
the user wants to store permanently individuals within the IEC system.

Now for the edges in this proposal are explained as follows:

\begin{itemize} 
\item The  LIKES edge represents a preference relationship that a user has with respect to an individual.  
\item Now the  PARENT edge represents an ancestor of the individual, in other words where the individual comes from.
\item Likewise the  HAS edge represents that users have a collection where can contain individuals.
\item Finally the KNOWS edge represents a relationship of friendship between users.
\end{itemize}

In figure \ref{fig:Nodes_Edges} illustrated the above mentioned, about nodes and edges.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/user_individual_collections.png}}
\caption{Nodes And Edges Representation.}
\label{fig:Nodes_Edges}       
\end{figure*}

In each vertex is necessary to store knowledge about the users, individuals and
also for the collections, this is what is known as property.

It is worth to mention that all vertices have a property called “element\_type”
among others. This property is used to identify which type of vertex is, for
example if vertex corresponds to a user then this property label the vertex as a
“person”. Likewise if the vertex corresponds to an individual this property
label it as an “individual” and is the same for any vertex we want to add to the
model. This property also has the purpose of being able to do operations on the
graph according to the vertex type.

The properties for the vertex “u” are the following:

\begin{itemize} 
\item id.  
\item created.
\item name.
\item element\_type.
\end{itemize}

The “id” property defines that the vertex u has a unique identifier in the set
of vertices, which means that there will not be duplicate vertices. The property
“created” defines the creation date of the vertex $u$. The property “name”
defines the name of the vertex u. The property “element\_type” defines the
element type that will be the vertex $u$.

In figure \ref{fig:User_node} the properties of vertex $u$  are shown.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_node.png}}
\caption{User Properties.}
\label{fig:User_node}       
\end{figure*}

The properties for the vertex $i$ are the following:
\begin{itemize} 
\item id. 
\item created. 
\item element\_type.
\item chromosome.
\item views.
\end{itemize}

The “id” property defines that the vertex $i$ has a unique identifier in the set
of vertices, which means that there will not be duplicate vertices. The
“chromosome” property defines the chromosome of the individual representing the
vertex $i$. which as defined above in this section. The  property “views” defines
the amount that users have seen this vertex $i$ The property “element\_type”
defines the element type that will be the vertex $i$. The property “created”
defines the creation date of the vertex $i$. % Is there a point in
                                % describing these properties? - JJ

In figure \ref{fig:Individual_node} the vertex $i$ properties are shown.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/individual_node.png}}
\caption{Individual Properties.}
\label{fig:Individual_node}       
\end{figure*}

The properties for the vertex “c” are the following:

\begin{itemize} 
\item id. 
\item element\_type.
\item created. 
\item name.
\end{itemize}

The “id” property defines that the vertex $c$ has a unique identifier in the set
of vertices, which means that there will not be duplicate vertices. The property
“element\_type” defines the element type that will be the vertex $c$. The
property “name” defines the name of the vertex $c$. The property “created”
defines the creation date of the vertex $c$.

In figure \ref{fig:Collection_node} the vertex $c$ properties are shown.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/collection_node.png}}
\caption{Collection Properties.}
\label{fig:Collection_node}       
\end{figure*}

In the same way that the vertices have properties also the edges, these
properties are defined as follows.

The “LIKES” edge has the following properties:

\begin{itemize} 
\item id. 
\item created. 
\item rate.
\end{itemize}

The “id” property defines that the “LIKES” edge  has a unique identifier in the
set of edges, which means that there will not be duplicate edges. The property
“created” defines the date of creation of the this edge. The property “rate”
defines the rate that the users give to the individual store in this edge.

In figure \ref{fig:Likes_edge} the edge “LIKES” properties and the relationships with the vertices $u$ and $i$ are shown.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_likes.png}}
\caption{LIKES edge Properties.}
\label{fig:Likes_edge}       
\end{figure*}

The “KNOWS” edge has the following properties.
\begin{itemize} 
\item id. 
\item created. 
\end{itemize}

The “id” property defines that the edge “KNOWS” has a unique identifier in the
set of edges, which means that there will not be duplicate edges. The property
“created” defines the date of creation of the edge.

In figure \ref{fig:Knows_edge} the edge “KNOWS” properties and the relationships between vertex “u”
are shown.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_knows.png}}
\caption{KNOWS edge Properties.}
\label{fig:Knows_edge}       
\end{figure*}

The “PARENT” edge has the following properties, and represent the parents of a new individual.
\begin{itemize} 
\item id. 
\item created. 
\end{itemize}

The “id” property defines that the edge “PARENT” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
The property “created” defines the date of creation of the edge.

In figure \ref{fig:Parent_edge} the edge “PARENT” properties and the relationship between vertex $i$ are shown.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_parent.png}}
\caption{PARENT edge Properties.}
\label{fig:Parent_edge}       
\end{figure*}

The edge “HAS” has the following properties, and represent the parents of a new individual.
\begin{itemize} 
\item id. 
\item created. 
\end{itemize}

The “id” property defines that the edge “HAS” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
The property “created” defines the date of creation of the edge.

In figure \ref{fig:Has_edge} the edge “HAS” properties and the relationship between vertices $u$, $i$ and $c$ are shown.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_has.png}}
\caption{HAS edge Properties.}
\label{fig:Has_edge}       
\end{figure*}

\begin{equation*}\label{eq:graphRelDef} 
\displaystyle 
\begin{split} 
V &= \{[u_1,u_2,u_3,...,u_n],[i_1,i_2,i_3,...,i_n],[c_1,c_2,c_3,...,c_n]\},\\ 
E&= \{[l_1,l_2,l_3,..,l_n],[p_1,p_2,p_3,...,p_n],[h_1,h_2,h_3,...,h_n],[k_1,k_2,k_3,...,k_n]\}\\ 
\end{split} 
\end{equation*}

Once  the vertices and edges have been defined a graph-based user-individual model is
proposed where  user vertices are related to the individual vertices, likewise
the vertices individuals are related to themselves when evolution exist between
them. % Grammar has to be improved - JJ
In that sense the vertices users and individuals are related to the
vertices collections as a vertex user can be related to a vertex collections and
a vertex collections collections is related to $n$ individuals vertex. In Figure
\ref{fig:user_moder} shows an example of the graph-based user-individual model.

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=8cm,height=8cm,keepaspectratio]{img/model_representation.png}}
\caption{Graph-based user-individual model.}
\label{fig:user_moder}       
\end{figure*}