### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Eliminating comments on the top

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi

% *** MATH PACKAGES ***
%
\usepackage{amsmath}

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}

% correct bad hyphenation here
\hyphenation{}

\begin{document}
% Title ideas here:

\title{Exploiting the Graph: Increasing Engagement in Collaborative Interactive Evolution}

\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678--2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Interactive evolution is the leading technique employed by 
Evolutionary Art researchers, where users' preferences guide the search. A common
practice is to develop Interactive Evolutionary Computation (IEC) algorithms
as web applications in order to give users easy access the application as they
often depend on volunteers who visit the system to interactively help with fitness assignement. 
There are several drawbacks that limit user participation 
such as human fatigue and boredom resulting from evaluating 
a large number of individuals. 
To tackle these issues, we propose the design of IEC systems using a human-centered 
approach, with a framework consisting of a social network of 
volunteers, interacting with a population consisting also of a network 
of individuals. In this work we suggest the use a graph model as a practical 
and efficient tool for mapping the relationships 
found in the system. A case study is presented as proof-of-concept, 
providing both conceptual and implementation details of a graph based model 
as it is  applied in the implementation of an IEC system. 
As result show, the data model was successfully applied in a gamification technique 
developed to increase user engagement.

%and novelty. % These two last issues are not too clear, maybe they
             % could be explained in that motivation.

\end{abstract}

% no keywords
\IEEEpeerreviewmaketitle

\section{Introduction}
Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
fitness evaluations are performed by persons within an interactive 
system.  Thus, the main loop of the EA needs the intervention of a
human to perform the quality assessment of each proposed solution.
IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}. 
Several IECs have been developed as web based systems \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users.
These systems have several inherent drawbacks arising from the very nature of 
the algorithms, namely, the human fatigue caused by the interaction, and
the boredom arising when users evaluate a large number of individuals 
many of which are not interesting or are very similar to each other.
IEAs also share some of the issues commonly found in volunteer based systems
\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
and the need to build trust between participants and application providers. 
For the provider there is also the difficulty of establishing 
the amount of time and resources
a volunteer is willing to spend on the system, and how they decide if they
participate or not \cite{JJ:2016}. In order to increase participation and 
engagement, and to tackle the issues mentioned above,  
IEAs must follow a human centered design, giving extensive attention 
to volunteer users, not only because their
explicit evaluation is essential, but also because the context of the 
interaction affects the system as a whole. Users interact not only with the graphical
interface, and each individual\'s phenotype, there is also the interaction
with other users, their previous evaluations and their own stream of activities
and past experiences. Thus users in this proposal are defined as dynamic entities 
interacting with an evolutionary computation having the fundamental purpose 
of evaluating individuals according to their preferences,
both explicitly or implicitly. In this work we propose giving the same
importance to users and their interactions as the population of 
individuals have in a traditional EA, even to the extend as using this
knowledge as an integral part of the evolutionary algorithm. We argue also
that a graph model is a viable representation for this purpose,  
presenting a natural way of mapping the relationships found in the system. 
Most IEAs have the following many-to-many relations:

\begin{itemize}

  \item {\bf Individual-Individual} The relationships between individuals in the
  population are naturally expressed as a tree, or a graph when both ancestors 
  and descendants are considered. Other semantic relations can also be of interest
  for instance similarity or composition.

  \item {\bf User-Individual}
  Many kind of relationships between users and individuals,
  can be expressed using common actions found in social networks;
  for instance: share, like, rate or add to collection \cite{Prodromou:16:AS}. In IEAs these
  relationships are used to assign a single or a collection of
  fitness values to individuals \cite{garcia2013evospace}.

  \item {\bf User-User}
  Relationship between users can also be modeled 
  as a social network, with well established semantics, algorithms and metrics \cite{ahuja1993network}.
  A graph model could enable researchers to find other ways of identifying leaders of 
  opinion or measuring the similarity between user's preferences. 
  These measures can then be used by recommender algorithms selecting 
  individuals according users' preferences.

\end{itemize}

These relations are naturally mapped to graphs, while other actions and 
properties can also be modeled as a semantic network using
basic subject–-predicate-–object expressions \cite{Prud'hommeaux:14:RT}.
Another reason for using a graph model is the tools available for
implementing and query the model, and the increased performance of
native graph database systems \cite{holzschuher2013performance,holzschuher2016querying}.

% Moving to related work - Mario
%\cite{Performance for the masses and Christian's previous work},

 % Can it be quantified? Maybe a reference to NodIO? - JJ
%Some authors have already tried to address this basic problem
%\cite{Frade:2010:EvoGAMES}. %Non sequitur. Is it something we are
                            %going to do here? - jj
%% Moving this to Related Work Section -Mario
% That section is not there yet. Forgot to paste? - JJ

A case study is presented as a proof of concept, providing both
conceptual and implementation details of a graph data model applied in the 
implementation of an IEA. The experiment was implemented using
the EvoSpace-Interactive framework, and consisted of a web application
for evolving artistic drawings. The graph database is then used to
measure the degree of relationship between users and their participation.

The remainder of this paper is organized as follows.
Section \ref{sec:interactive} presents related work on the topic 
of Interactive Evolution.
Then, Section \ref{sec:evospace-i} presents the computational platform on which 
the main proposal of this work is developed. The proposed graph model is 
presented in Section \ref{sec:graph}.
The experimental work and results are presented in Section \ref{sec:experiments}.
Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

\section{Related Work}
\label{sec:interactive}
% IEC 
% Collaborative
% Users - Volunteers - Engagement
% Too much data
% Relational MySQL, Key-Value, Mongo, Redis, Memcached
% Not well suited for certain relations.
% Graph DBs, Performance, Query Langs 
% Are they any good? 
% Titan, Neo4J

The IEC the approach has been used in a wide variety of applications and problem domains
and remains an active area of research. However, in this work, emphasis is given to IEAs 
implemented as volunteer based collaborative systems; where volunteers interact and evaluate 
an evolving population, thus guiding the search based on an aggregate of their subjective 
preferences. What follows is not intended as a comprehensive survey, only a review of 
the most relevant contributions to the present work, with special attention put in
data structures and data management requirements.

From the earlier works in the IEC field the notion of user collaboration in a natural way
has been proposed. A notable application was the Galapagos Project \cite{sims1997interactivity},
an exhibit in the Tokyo Multimedia Museum (1997--2000) were visitors interacted with images presented in 
twelve displays by selecting those they find most aesthetically interesting by standing on
step sensors in front of them. An early example of web based collaboration is the work by 
Langdon \cite{langdon:2004}, evolving fractal representations of virtual creatures. It proposed a 
distributed EA using a global population stored on a central web server sending 
portions of the population to remote clients using JavaScript. Users evaluated individuals locally 
and returned to the server those they liked more, to be distributed over the web.
Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} use web-based IEAs 
to evolve artistic artifacts using a generative encoding, compositional pattern producing networks.

In all these systems, the interactive process is captured by the system. The genetic lineage of each
individual is presented in a graphical interface 
where users can visualize the relationships between them, and evolutionary operators used in
their creation; they also display information regarding users and their collaborations.  Common offerings show 
interesting individuals ordered by categories, number of votes, and best of the week among others. In a recent work by
Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
new designs of robot locomotiom. Collaboration is encoureged by gamifying the system 
using the maximum distance indicator to inspire the user to try and “beat” previous designs.

Implementing recommendation algorithms or adding gamification techniques have the goal of encouraging participation 
but require the use of a database system for managing the additional data required. 
For example Picbreeder \cite{picbreeder} is implemented with a service oriented
architecture using a relational database and the file system for genotypes and images. Other works employ the 
use of No SQL databases like the work of Seyama and Munetomo \cite{seyama2016development} where they
employ a Platform as-a Service (PaaS) and NoSQL database to facilitate the collaboration 
among a large number of users for the interactive modelling of 3D glasess.  This system reduces the user fatigue
by using a collaborative filtering algorithm to show only the information utilized by similar users.

Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common in current applications
like social networks. In order to deal with large datasets of connected data found in these systems,
graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
In this work the use of a graph data model for representing the interactive process of C-IEC systems is proposed.
Moreover, an proof-of-concept implemantation of an IEA incorporating the graph model is presented.

%Finally, a comparasion of between three deployments of the system, with different increase the collaboration of volunteers.

\section{EvoSpace-Interactive}
\label{sec:evospace-i}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/UI_ed01.png}
    \caption{User interface of an EvoSpace-Interactive application.}
    \label{fig:web}
\end{figure}

As a case study, a IEA was developed using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. %citation
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

\subsection{Fitness Assignment}
\label{sec:assignment}
The ES-I platform employs a collaborative technique,
where several registered users assign a quality assessment to a single
individual and then an aggregated fitness value is calculated. The fitness
assigned to each individual depends on the taste of each particular user, 
resulting in a many-to-many relationship between users and individuals. 
Many systems query this user-individual relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular individual, or the the user with more participation 
\cite{picbreeder}.
In order to do this, data about individuals 
must be permanently stored, even
if they are no longer in the active population.

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those individuals they like, or by taking individuals
from their friend\'s collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, an individual sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the individual using
a five star rating measure or
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder to store those individuals a user prefers and wishes
to save. After the user finishes interacting with the individual
on the Wall, he can choose to retrieve a new one from the population.
At the left side, is the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, a user can browse the content of each collection and from
there share images through the social network.
When a user browses over an individual a detail pane shows how many users have
liked the individual. The pane also includes a link to the individual's 
details, the parents, genetic operators that created it, and genealogy information.

\begin{figure*}[!t]
    \centering
        \includegraphics[width=3.5in]{img/framework.png}
    \caption{IEC Human-centered framework.}
    \label{fig:uc_framework}
\end{figure*}

%%
%% TO DO: Talk about the Model (in the MVC sense) of this system: PostgreSQL, Redis, 
%%

 \section{IEC Human-centered framework}
\label{sec:graph}
The general goal of this research is to develop a human-centered \cite{gasson2003human} framework
for interactive evolutionary computation (IEC) in order to increase
participation and also to minimize the amount of evaluations needed for the
evolutionary process in given IEC application. We deliberately use the 
more general
term Human instead of User, because humans are not always playing the role of
users in software engineering sense, they could just be interacting 
naturally with their environment and with out knowing they are implicitly
evaluating a possible solution. In other cases the evaluation could be done
by the natural interactions of a group of humans in a collaborative fashion
or not. For instance the IEC algorithm could by trying to optimize 
the design of an exit sign in an airport, then the quality of this sign could be
determined by the average the time it takes a sample of people to 
find the exit. In this example there could be certain collaboration
between users if they are not traveling alone, of course this is only one of
the considerations designers could make.

The human centered framework consists of three high level components depicted
in figure \ref{fig:uc_framework}:

\begin{itemize}
  \item {\bf Interactive Evolutionary Computational system} 
  This is the real world system that we are going to represent in the model, 
  it consist of human users and their interactions with one or more individuals
  from
  the population. There are many ways in which humans could interact 
  with individuals, this could be through a mobile device, a user interface 
  or by interacting with real world objects generated as part of the algorithm 
  \cite{de2014artists,de2013unplugging}. 
  There is also the possibility that fitness or even part of the search 
  to be done by devices lent by the human \cite{merelo2016performance}.
  The information gathered through the interaction is the primary concern
  as it will guide the search.

  \item {\bf IEC Data Model}
  A data model is used to describe the IEC system prior to a physical 
  implementation. 
  %% This block is maybe not needed but in our ES-I implementation
  %% we use Relational (PosgreSQL), Objects (JSON-Redis) and Graphs (Neo4J)
  Depending on the domain several techniques can be used.
  Parts of the system could be better described by an Entity relationship 
  modeling to be used in a relational database or a Class Digram for a 
  key-value store. 
  %%
  %%
  A graph is proposed for modeling the social network
  of users and their interaction with the population. This database 
  is the back-end of the system, is used by the EA and the user interface.

  \item {\bf Evolutionary Algorithm} 
  The EA algorithm queries the graph model, and then updates those nodes that 
  are part of the population. The EA could also be done with the help of
  humans, for instance in the XY project \cite{de2013unplugging} artists physically created
  the new population.  
\end{itemize}

\section{IEC Graph Data Model}

A graph is composed of two fundamental entities: nodes and relationships,
both having attributes. Nodes are abstractions of the objects in the real 
world system, and relationships describe their interactions. Types describe
the attributes describing an entity, First we describe the types of nodes and
relationships in the system.

\subsection{Node Types}
\begin{itemize}

\item {\bf Users} This nodes represent the human users of the system and have 
the attributes: {\tt id} a unique identifier, {\tt name}  and {\tt created}
specifying the  creation date (time stamp) of the node.

\item {\bf Individual} This type of node has the following attributes: 
{\tt id}, {\tt chromosome}  and {\tt views} this is 
the amount that users have seen or interacted with this individual and
 {\tt created}.

\item {\bf Collection} This type of nodes represent the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if this collection
contents can be are shared with others. 
\end{itemize}

\subsection{Relationship Types}

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
an individual in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
an individual multiple times, so the date of the interactions is also of
interest.

\item {\bf Knows} The relation connects two users that know each
other. This relation is "unique" meaning that it can not be repeated between a pair of nodes.

\item {\bf Parent} Describes what individual is the parent of a new individual and is also unique.

\item {\bf Has} The relation describes an ownership relation between users and
those collections they own. Also a collection {\tt has}  individuals stored
in them.  In figure \ref{fig:has} the {\tt has} relationship 
between certain nodes $user$, $individual$ and $collection$ is shown.

\item {\bf Shared} When a user takes an individual from a friends collection, this interactions 
is described by a {\tt share} relationship, between the user and the individual.

\item {\bf Saved} When a user takes an individual from the wall interface, this 
interactions is described by a {\tt save} relationship. 
\end{itemize}

\begin{figure}[!t]
    \centering
        \includegraphics[width=2.5in]{img/edge_properties_has.png}
    \caption{ The {\tt has} relationship type example and attributes.}
    \label{fig:has}
\end{figure}

Is important to notice that this model is still partial view of the IEC system,
and other models can could be used in practice when implementing the system. 
For instance a relational database can be used for the implementation of a
user profile and authentication module. Also a copy or cache of the population 
could be kept in a key-value store in memory for performance reasons. 
Nevertheless, the model can give important insights of the behavior of the 
system.

\section{Case Study}
\label{sec:experiments}

We present in this section the results of a comparison between three versions
of the same IEC application. The goal of the experiment is to compare two 
gamification techniques to increase volunteer participation against a third 
without any thechnique applied.

The main difference between both versions employing gamification is how the fitness of individuals
is calculated and the information presented to users.

\subsection{Gamification techniques}
\label{sec:gamification}

Gamification is defined by Deterding et al. as
``the use of game design elements in non-game contexts'' \cite{deterding2011game}.
The gamification element employed in this work is rewarding mechanism 
\cite{dubois2013understanding}, in general they consisting of a reputation systems with score points, 
levels and leader boards. Points are awarded to users in response of
the accomplishment of certain activities that need to be encouraged. Levels depend
on the score and certain features of the game are only available to gamers when 
they reach a giving level.

The rewarding mechanism as it is applied in these versions 
consists in giving more importance to the preference of those users having a higher reputation
as given by the score points and experience levels.

The score achieved by users depends on the actions he does in the system as
each time a user does one of these actions their score is incremented by one:
\begin{itemize}
\item Start a session.
\item Rate an individual.
\item Create a collection.
\item Save an individual of the wall to a collection.
\item Save an individual from a friend's collection.
\item Explore collections of other friends.
\end{itemize}

Only the ``Start a session'' and ``Explore collections of other friends'' actions 
are not found in the graph model as these are stored in the PostgreSQL relational
database system.

Two scores are used to determine the weight of a user's preference:
\begin{itemize}
\item Experience. This value depends on the score and is a value 
between 0 and 100. 
% More details needed? Is done by a Fuzzy Inference System but I think
% the description would take to much space.

\item Participation. This value is simply the degree of the user node 
in the graph. This measure also considers the number of users he knows,
perhaps some of the users where invited to the system by him.   
\end{itemize}

Three versions were compared:
\begin{itemize}
\item Base (B). All users have the same weight.
\item Non Graph Gamification (G). Only experience is considered
\item Graph Gamification (GG) . Both experience and participation are considered.
\end{itemize}
When gamification was employed, all score values known where presented to users
and a ranking of users by weight was shown in a window, a screen capture of
this interface is shown in figure \ref{fig:top-users}.

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/users_known_1.png}
    \caption{Comparasion of two users showing the {\tt knows} relationship.}
    \label{fig:users-graph}
\end{figure}

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/leaderBoard.png}
    \caption{User interface showing top users.}
    \label{fig:top-users}
\end{figure}

\subsection{Set-up}
\label{sec:setup}

The three versions of the EvoDrawing web application were deployed to the
Heroku cloud service with the same configuration: using the
Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB and RedisTOGO plug-ins.
Table \ref{tab:params} shows the parameters used for the evolutionary algorithm. The graph database 
system used in the implementation was Neo4J as it was allready available as a plug-in in Heroku, is well documented
and a scalable solution \cite{miller2013graph,holzschuher2013performance}.

\begin{table}
  \small
  \caption{ Parameters for experiments.  }
  \label{tab:params} 
  \centering
  \small
  \begin{tabular}{l  c   }
    \hline\noalign{\smallskip}
     Parameter & Value \\
    \noalign{\smallskip}\hline\noalign{\smallskip}
    Initial Population Size   & 80 \\ \hline
    Sample Size & 1 \\ \hline
    Step Size & 8 Samples \\ \hline
    Mutation &  \\ \hline
    Selection & Tournament \\ \hline
    Tournament Size &  6 \\ \hline
  \end{tabular}
\end{table}

% Anuncements
At the start of each of the experiments, a call to participation was promoted through social networks
particularly on Facebook and Twitter social networks.
Is important to note that participants needed to log-in to facebook and give permissions to the application
in order to participate, there were no anonymous volunteers. This permited the automatic access to
each participant's social network and to store a local copy of those friends also participating.

The link used in the call for participation
was shortened URL, in order to obtain additional data and analytic of the visitors.
In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,
along with the short URL and the analytics link. In the same table a link to the GitHub 
application repository used to deploy to Heroku is also listed.

\begin{table*}
  \small
  \caption{ URLs used for each deployment.}
  \label{tab:urls} 
  \centering
  \small
  \begin{tabular}{l l  l l l}
    \hline\noalign{\smallskip}
     Deployment & Heroku URL & Shortened URL & Analytics & Github Repository \\
    \noalign{\smallskip}\hline\noalign{\smallskip}
    B   & evodrawings01.herokuapp.com & goo.gl/jLis4Q & goo.gl/jLis4Q.info &  \\ \hline
    G   & evodrawings02.herokuapp.com & goo.gl/jqjNy5 & goo.gl/jqjNy5.info &  \\ \hline
    GG  & evodrawings03.herokuapp.com & goo.gl/J8TCe1 & goo.gl/J8TCe1.info &  \\ \hline
    \end{tabular}
\end{table*}

Only data for the first week of deployment was considered for the experiments, and they where conducted 
between January and May of 2016.

% Time
\subsection{Results}
\label{sec:results}

Before launch, each deployment was first tried with a few beta testers. When applying the gammification
technique for the first time a problem was found in this beta stage. Some users were chating by 
giving a rating to an animation even before it was returned from the server, this was done by just
constantly pressing the mouse button. This is a common problem found in systmes appling leader boards.
By making the rewards for tasks achievements visible to other players or providing leader boards are ways of 
encouraging players to compete \cite{hickman2010total}. Because the  problematic consequences of competition, 
which can result in negative conduct, low cooperation and collaboration, or disadvantaging certain player
demographics such as women \cite{kumar2013gamification}, best-practice gamification designs try to
refrain from using this element. In order to measure the effect of the leader board in the system, the 
problem of cheating was mitigated by disabling the buttons for a certain amount of time. But the problem 
persists, competitive users will try to get in to the leader board by other means. 
% 
% Future work could compare the use of gamification with out leaderboards
%

The results of each of the three experiments in terms of participation are detailed next.

\subsubsection{ Base deployment}
After a week of the anouncement the total number of volunteers for this experiment was 53 registered users. 
The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
experiment a total of 500 new individuals where generated. The total number of evaluated
individuals for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a blue line.

\subsubsection{ Non Graph Gamification deployment}
The total number of volunteers for this experiment was very close to those obtained in the previous
experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
with a total of 2596 relationships. These numbers are close to those obtained in the base experiment,
with a marginal increase in participation. The increased participation also helped in the creation of
more individuals for a total of 648. The total number of evaluated
individuals for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a red line.

\subsubsection{ Graph Gamification deployment}
This experiment had the higher number of volunteers for for a total of 68 registered users. 
In this experiment a total of 3594 new individuals where generated in the experiment. 
The total number of evaluated individuals for each volunteer is presented in figure 
\ref{fig:top-ranked-participation} in green.

The results presented in figure \ref{fig:top-ranked-participation}, show that those users with
higher level of participation in deployments B and G, had about the same behavior, but the difference
came with users with medium level of participation, in this case participants where more engaged in deployment
G. When comparing all the experiments the deployment GG had the higher number of participation in users
of all levels.

When comparing the networks of users, there is also a difference shown between experiments.

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/user_known_2.png}
    \caption{User network view of the Base deployment}
    \label{fig:B-network}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/user_known_3.png}
    \caption{User network view of the Graph Gamification deployment}
    \label{fig:GG-network}
\end{figure}
\begin{figure*}[!t]
    \centering
        \includegraphics[width=5.5in]{img/comparison.png}
    \caption{Top ranked number individuals rated by user using a
     logarithmic scale.}
    \label{fig:top-ranked-participation}
\end{figure*}

Using the Neo4J graph database and the Cypher query language when monitoring the experiments provided 
views on the connected data difficult to replicate in a relational database. As an example, when comparing the
deployments B and GG based on their network of users (shown in figures \ref{fig:B-network} and \ref{fig:GG-network}
respectivly) additionaly to the differences in the number of nodes, certain patterns can also be observed.

\section{Conclusion}
\label{sec:conclusions}

\section*{Acknowledgment}
Space reserved for Acknowledgment text, Lorem ipsum dolor sit amet, consectetur adipiscing elit,
sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis 
nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor
in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\bibliographystyle{IEEEtran}
\bibliography{evospace-i,volunteer,biblio}

\end{document}

\end{document}