### 2017-04-05 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Authors

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[GECCO '17]{the Genetic and Evolutionary Computation Conference 2017}{July 15--19, 2017}{Berlin, Germany}
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}

\begin{document}
\title{Modeling Human Interactions in Collaborative Interactive 
 Evolutionary Computation}

\author{Mario Garc\'ia-Valdez,\\ Jos\'e-C  Romero}
\affiliation{%
  \institution{Dept. of Graduate Studies,\\ Instituto
Tecnol\'ogico de Tijuana, Tijuana, M\'exico}
  }
\email{(mario|jcromero)tectijuana.edu.mx}

\author{Juan-J.~Merelo}
\affiliation{%
  \institution{Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain}
  }
\email{jmerelo@ugr.es}

\author{Francisco Fern\'andez-de-Vega}
\affiliation{%
  \institution{Grupo de Evoluci\'on Artificial,\\
Universidad de Extremadura, Spain}
  }
\email{fcofdez@unex.es}

\begin{abstract}
The necessary intervention of humans in interactive evolutionary
computational systems has inherent drawbacks arising from the very nature of 
the algorithms, namely, the human fatigue caused by the interaction, and
the boredom arising when users evaluate a large number of artifacts.
To tackle these issues, in this paper we
propose a human-centered framework to model complex interactions on these systems.
A case study is presented where the model is applied in the 
development of a collaborative evolutionary interactive system. 
% Both conceptual and implementation details are provided where the   
% technique is used to measure and increase user engagement and participation.
% Our experiments show that the model can be successfully applied in a
% gamification technique developed to increase user engagement, which
% implies that this technique can successfully be used to 
% decrease user fatigue and boredom, and thus increase the {\em performance} of the
% interactive system.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003120.10003130.10003233</concept_id>
<concept_desc>Human-centered computing~Collaborative and social computing systems and tools</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010178.10010205.10010206</concept_id>
<concept_desc>Computing methodologies~Heuristic function construction</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010178.10010205.10010207</concept_id>
<concept_desc>Computing methodologies~Bio Inspired Algorithms</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Human-centered computing~Collaborative and social computing systems and tools}
\ccsdesc[300]{Computing methodologies~Heuristic function construction}
% We no longer use \terms command
%\terms{Theory}

\keywords{Interactive evolutionary computation, Human Centered Computing}

\maketitle

\input{paper-body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sigproc,../bib/biblio,../bib/evospace-i,../bib/volunteer,../bib/geneura}

\end{document}