### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Small changes in abs

      \documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[GECCO '17]{the Genetic and Evolutionary Computation Conference 2017}{July 15--19, 2017}{Berlin, Germany}
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}

\begin{document}
\title{Modeling Human Interactions in Collaborative Interactive 
 Evolutionary Computation}

\author{Mario Garc\'ia-Valdez, Jos\'e-C Romero}
\affiliation{%
  \institution{Dept. of Graduate Studies,\\ Instituto
Tecnol\'ogico de Tijuana, Tijuana, M\'exico}
  }
\email{mariosky@gmail.com}
\email{josec.romero@tectijuana.edu.mx}

\author{Juan-J.~Merelo}
\affiliation{%
  \institution{Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain}
  }
\email{jmerelo@ugr.es}

\author{Francisco Fern\'andez-de-Vega}
\affiliation{%
  \institution{Grupo de Evoluci\'on Artificial,\\
Universidad de Extremadura, Spain}
  }
\email{fcofdez@unex.es}

\begin{abstract} The necessary intervention of humans in interactive
evolutionary computational systems has inherent drawbacks arising from the
very nature of the algorithms, namely, the human fatigue caused by the
interaction and the boredom arising when users evaluate a large number of
artifacts. To tackle these issues, in this paper we propose a human-centered
framework that can be used to increase volunteer participation in collaborative interactive evolutionary computational (C-IEC)
systems by using gamification techniques. A case study is
presented where the model is applied in the development of a collaborative
evolutionary interactive system.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003120.10003130.10003233</concept_id>
<concept_desc>Human-centered computing~Collaborative and social computing systems and tools</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Human-centered computing~Collaborative and social computing systems and tools}
\keywords{Interactive evolutionary computation, Human Centered Computing}

\maketitle

\input{paper-body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sigproc,../bib/biblio,../bib/evospace-i,../bib/volunteer,../bib/geneura}

\end{document}