### 2017-11-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-body.tex
old uncommitted changes

\section{Introduction}

Interactive Evolutionary Computation (IEC) systems include persons in
a computational system performing different kind of tasks, which brings particular challenges
to their designers: human {\em computational cycles} are scarce, slow
and expensive to acquire, there is human 
fatigue caused by the interaction \cite{ie1}; besides, boredom arises
when users evaluate a large number of phenotypes. In this paper we
intend to solve these issues by developing a  human-centered 
\cite{gasson2003human} software framework that can be used to 
increase volunteer participation in C-IEC systems by using
gamification techniques, which are used to {\em human computational
  cycles} more abundant and possibly decrease boredom.

Gamification is the use of game design elements in non-game contexts \cite{deterding2011game}. 
The gamification element employed in this work is a rewarding mechanism
\cite{dubois2013understanding}. An example of the use of this technique is a recent
work by Wagy \& Bongard \cite{wagy2014collective} in which interaction is
needed for developing new designs of robot locomotion.
Collaboration is encouraged by gamifying the system using the maximum distance
indicator to inspire the user to try and ``beat'' previous designs. Also 
Seyama \& Munetomo \cite{seyama2016development} propose the reduction of user
fatigue by using a collaborative filtering algorithm to show only the
information utilized by similar users. In Section \ref{sec:framework}
we present our proposed framework. Next, the EvoDrawings application case study 
is presented in Section \ref{sec:evodraw} and the
results are discussed in Section \ref{sec:results}. Finally, some concluding
remarks are provided in Section \ref{sec:conclusions}.

\section{Human-Centered C-IEC Framework}
\label{sec:framework} 
The main design considerations of the framework are explained next: 
\begin{itemize}
\item {\bf Users are volunteers.} Users donate their computing resources, so they are 
unaccountable and sometimes they try to game the system. Project owners must actively promote and
design the interactive system to engage volunteers \cite{oh2015clicking}. % Define
\item {\bf Users are not alone}
  Relationships between users in an interactive evolutionary algorithm can be modeled
  as a social network, with well established semantics, algorithms and metrics 
  \cite{ahuja1993network}. 
\item {\bf Interaction is a stream of actions.}
  Real time processing of users' actions could be needed for certain applications when data is 
  captured by sensors, or directly captured as user input. For example, social networks encourage
  users to publish their interactions with other users, media objects and places.
  there are initiatives like the W3C Activity Streams 2.0  specification, used for 
  representing common activities in social web applications \cite{json:streams}. 
\end{itemize}

A graph is proposed for modeling the social network of users and their interactions 
with candidate solutions and the relationships between them in the population.
The graph database system used in the implementation is Neo4J, which is a scalable solution 
\cite{miller2013graph,holzschuher2013performance}. 
This graph is also used to increase engagement through gamification, as explained in the next section.

\section{Case Study: EvoDrawings Gamification}
\label{sec:evodraw}
As a case study, a C-IEC application was developed by extending the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}, 
with the open source code hosted in GitHub\footnote{\url{https://github.com/mariosky/evo-drawings}}.
The rewarding 
mechanism as it is applied in EvoDrawings gives more importance 
to the preference of those users with higher reputation
as given by their score points and experience levels.  
Each time a user does one of these actions their score is incremented by one:
start a session, rate a phenotype, create a collection, save a phenotype of 
the wall to a collection, save a phenotype from a friend's collection, and
explore collections of other friends. Each of these actions are stored in
the Neo4J graph. Two variables are used to determine the weight of a user's 
preference:

\begin{itemize}
\item {\bf Experience}: This variable depends on the score and is a value 
between 0 and 100. A new user starts at zero, and the experience increases until
it reaches 100 actions.

\item {\bf Participation}: This variable is simply the degree of the user node 
in the graph (number of edges).    
\end{itemize}

Three versions of EvoDrawings were compared: Base (B): All users have the same weight;
Non Graph Gamification (G): Only experience is considered; and Graph Gamification (GG): 
Both experience and participation are considered.

\section{Results}
\label{sec:results} 
Before release, the deployment was first tried with a few beta testers. 
When applying the leader board gamification technique for the first time, a 
problem was found: some users were cheating by giving a
rating to an animation even before it was returned from the server, this was done by just
constantly clicking the mouse button. Cheating is a common problem found in systems using leader
boards because by making the scores visible to other players they are encouraged 
to compete against each other \cite{hickman2010total}. The version used in experiments disabled the button until 
the drawing animation was over. The results of each of the three experiments in 
terms of participation are detailed next.

\begin{table}
  \small
  \caption{After a week of the announcement the total number of volunteers, 
  nodes and edges in the graph and analytics URLs}
  \label{tab:urls} 
  \centering
  \small
  \begin{tabular}{l l l l l}
    \hline\noalign{\smallskip}
     Deployment &  Users &  Nodes &  Edges \\
    \noalign{\smallskip}\hline\noalign{\smallskip}
    B   & 53 &  595   & 2220   \\ \hline
    G   & 54 &  648   & 2596   \\ \hline
    GG  & 68 &  932   & 3594   \\ \hline
    \end{tabular}
\end{table}

Table \ref{tab:urls} shows the total number of volunteers, nodes and edges 
in the graph after each experiment. 
% Moreover, the total number of evaluated 
% phenotypes for each volunteer is presented in figure 
% \ref{fig:top-ranked-participation} where users are ranked by the 
% number of phenotypes they rated. 
% Same as in the other, figure missing - JJ
When comparing all the experiments the GG deployment had the higher
number of participation, besides attracting also the higher number of users.

\section{Conclusions}
\label{sec:conclusions}

In concordance with the results obtained in other 
browser-based volunteer systems, after 
applying the gamification techniques, user participation was
increased. By the nature of the experiment, we will have to repeat it
several times and check causal correlation, but in general all
measurements indicate that is the case.

One of the interesting future lines of work would be to look a bit
more closely at the behavior of users as they are rating artifacts 
in the web system. These initial experiments hint at a possible power law, which might indicate that
the IEC system could be self-organizing, a process that would allow it
to reach a critical state, as has been found in software repositories,
for instance \cite{Merelo2016:repomining}. It would be also
interesting to study the possible negative effects of using   
gamification techniques to improve engagement, like cheating or
literally {\em gaming} the system to defeat competition. 
Finally, the refinement of the proposed Human-Centered framework will need
more case studies and further multi-disciplinary research.

\begin{acks}
This work has been supported in part by: de Ministerio espa\~{n}ol de Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-\{1,2,3\}-P (UGR-EPHEMECH), by Consejo Nacional de Ciencia y Tecnolog\'{\i}a under project 220590, and by Junta de Extremadura FEDER, project GR15068

\end{acks}