### 2018-01-11 2018-IPMU/Gamification/paper.tex
changes proposed my pablo

\documentclass{llncs}
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{url}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{subfigure}
\usepackage{amsmath}

\begin{document}
\sloppy

\title{Increasing Performance via Gamification in a Volunteer-Based Evolutionary Computation System}
\titlerunning{Increasing Performance via Gamification in a Volunteer
  EC System}

\author{Mario Garc\'ia-Valdez\inst{1} \and Juan J. Merelo Guerv\'os\inst{2} \and  Lucero Lara \inst{1} \and Pablo Garc\'ia-S\'anchez \inst{3}}

\institute{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico
\and
Universidad de Granada, Granada, Spain
\and
Universidad de C\'adiz, C\'adiz, Spain
\email{mario@tectijuana.edu.mx}\\
\email{jmerelo@geneura.ugr.es}}

\authorrunning{Garc\'ia-Valdez, Merelo, Lara \& Garc\'ia-S\'anchez}

\maketitle

\begin{abstract}

Distributed computing systems can be created using volunteers, users who spontaneously, after receiving an invitation, decide to
provide their own resources or storage to contribute to a common
effort. They can, for instance, run a script embedded in a web page;
thus, collaboration is straightforward, but also ephemeral, with
resources depending on the amount of time the user decides to
lend. This implies that the user has to be kept engaged so as to
obtain as many computing cycles as possible. In this paper, we analyze
a volunteer-based evolutionary computing system called  NodIO
with the objective of discovering design decisions that
encourage volunteer participation, thus increasing the overall
computing power. We present the results of an experiment in which a
gamification technique is applied by adding a leader-board showing the
top scores achieved by registered contributors. In NodIO,
volunteers can participate without creating an account, so one of the questions we wanted to address was if the
need to register would have a negative impact on user participation.
The experiment results show that even if only a small percentage of
users created an account, those participating in the competition
provided around 90\% of the work, thus effectively increasing the
performance of the overall system. %Pablo: how the performance is compared with respect to the previous version? Less time?
% In this case is more work, more indivudals put in to the
% pool or more generations - Mario

\keywords{Distributed Evolutionary Algorithms, Volunteer Computing,
  Socio-Technical Systems}
\end{abstract}

\section{Introduction}
The world wide web is an increasingly reliable and high-performance
operating system for running distributed applications. Besides the
maturity of HTTP, the underlying protocol itself, there are other
factors: the JavaScript virtual machine embedded in every browser
and the REST interface, providing a {\em de facto} standard interface for
communication. A distributed system based on the browser can be
easily created using these mechanisms, which lately have been enhanced
by WebSockets for lower latency; you can start to gather users just by
announcing the URL. This approach for creating distributed
experiments is called {\em volunteer}, {\em cycle-scavenging},
or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the Internet
\cite{david-seti:home}.

In this context, we are mainly interested in evolutionary algorithms
\cite{milani2004online,sherry2012flex,jj-ppsn98}%,
% population-based search and optimization methods inspired by
% the theory of Evolution and its molecular basis.
. The fact that they are population-based makes them suitable for a
straightforward distribution of work to different clients, and since
they have an asynchronous nature, there is no considerable impact on
their performance, and might even be beneficial for a number of
reasons, including an increased diversity
\cite{cantu-paz:migration-policies}. All these properties make them
ideal candidates for volunteer computing setups such as the one
presented in this paper.
%, at least if some minimal changes are made to
%the usual sequential or parallel algorithms that rely on a static
%setup.  -- Not needed, Mario

But even if adaptation is straightforward, the application of volunteer
computing techniques to the development of distributed evolutionary algorithms
still has several open issues. An important research aspect is
approaching the problem as a socio-technical system
\cite{vespignani2009predicting,merelo2015designing}, which integrates
user decisions and behavioral patterns in the model; this
includes trying to optimize the number of users or the overall time
they contribute in a particular
experiment. The challenge is to design a system that, whatever the
number of users available and willing to perform the experiment, it is able to
maximize their contribution to the evolutionary algorithm. One obvious
way of achieving that is trying to find as many users as possible. But
another approach, and the one that we will be following in this paper, is
to delve into the socio-technical nature of the system, using
gamification techniques to try and improve the amount of time every user will
lend to the system. However, there is a trade-off with gamification,
since users have to be identified in some way to participate, and thus
it has to include some way of user authentication,
which is a hurdle that might take users away; however, the open design %Pablo: use colon instead semicolon?
of the system we use means that we can actually mix anonymous (with no
gamification) and authenticated (gamified) users, with only the small overhead
of adding the gamification mechanism.

This trade-off between ease of use, allowing users to
participate by just visiting a web page, and customization, making
registration a requirement in order to % I just simplified the sentence -Mario
participate, is what we will analyze in this paper. In a socio-technical system such as this one, every
decision the user has to take has an influence on the overall
performance, and registration could be seen as a hurdle to volunteers
trying to participate in an ephemeral experiment, so it is a challenge
to analyze what actually happens in this case; quite clearly, the
inclusion of gamification techniques has an impact on page loading and
complexity of server-side programming; if the number of users
or the amount of time they contribute to the experiment is not enough,
it would be better to revert NodIO to a purely anonymous system.

The rest of the paper is organized as follows: The state of the art in opportunistic/volunteer distributed evolutionary
computation (EC) and Gamification Techinques as presented next. Section \ref{sec:experiments} will describe the
framework and the problem used in the experiments, which are publicly
available under a free license. The results of each of the steps in the incremental design are presented in Section
\label{sec:results}, to finally wrap up with the conclusions. %Pablo: check out this paragraph! Section 3 is not describing the problem and framework, it is only discussing on related work on gamification! I suggest combine section Gamification into Section 2, and separate Section 4.2 (Results) as a new Section before Conclusions.
% Done - Mario

\section{Related Work}
\label{sec:soa}

Volunteer computing involves users who decide to run a program that acts
as a client or as a peer in a distributed computing experiment
and, as such, has been deployed in many different ways since the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \cite{david-seti:home}, which used
a client that acted as an screensaver.%, or a
%high-throughput queuing system such as HTCondor \cite{HTCondor}.
However,
the dual facts of the introduction of JavaScript as a universal language for the
browser and the browser itself as both an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks in what is generally described as BBVC or Browser-Based
Volunteer Computing \cite{fabisiak2017browser}.

% ystems based on the
% JavaScript/browser combination
% emphasize the ephemerality, ease of use, and universality, while
% systems such as HTCondor or BOINC \cite{boinc_grid04} might be more adequate for work that
% requires higher availability of volunteer resources.

In order to attract volunteers, many systems rely on downloadable clients;
% , and in
% some cases like HTCondor, they also require administrative access to
% certain resources
Fabisiak \cite{fabisiak2017browser}
analyzes the different capabilities of Internet and browser-based
systems, including their ability to handle large sets of input data
and time-consuming problems, concluding that BBVC is mostly apt for
short, data-light tasks.

As indicated in the introduction, the human is an integral part of the system, which we can consider a {\em human computation} \cite{quinn2011human}. Giving the user more control, and adding visibility to their participation might have positive effects. In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' in which the name of the volunteer, the number of projects, GFLOPS, country and team are displayed. %Pablo: I suggest to move this paragraph above the previous one, as it works as a better introduction to this kind of systems and can be linked to the next paragraph.
% Moved - Mario

Several authors have already described systems using JavaScript either
for unwitting
\cite{unwitting-ec,boldrin2007distributed,apolonia2012enhancing} or volunteer
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed EC and it has been used by several
authors ever since. In
fact, \cite{pan2015gray} performs an analysis of what
is called {\em Gray computing} covering aspects of
feasibility, cost-effectiveness, change in the user\'s experience and the
architectural optimization needed, concluding that the computing power
available is vast and it can be cost-effective to use it.

% Recent works have been using crowdsourcing in order to train robots
% for human interaction, a representative work of this trend is that of
% Anetsberger \& Bongard \cite{anetsbergerrobots} in which they propose
% training simulated robots for the grounding of language symbols. They do this by using human observers that issue arbitrary commands to these robots via the web, providing positive or negative reinforcement to the resulting actions of the robot. For interaction, they use the Twitch streaming platform at which users can issue commands through the
% platform\'s integrated chat service. Subjects were incentivized to
% interact with the system by GUI features that provided participants
% with a sense of involvement with the simulation. This kind of system
% could add additional dynamics, because users can collaborate or
% even compete with each other when issuing their commands. This approach has been called {\em social cloud} \cite{6404452}, emphasizing the fact that user participation is gamified in order to obtain the most from it \cite{7027564}.

Many systems described above do not go any further than trying to find out how many users join the effort, and how many of them the system can support. In fact, systems such as the one described in
\cite{gecco07:workshop:dcor} had severe scaling issues; some of them
also tried to find out how much time was needed to find the solution
or, alternatively, how many users would be needed to be competitive
against single-user single-computer implementations of the same
algorithm. Lately, researchers have tried to integrate volunteer
computing techniques as a part of a larger distributed evolutionary computing
effort \cite{leclerc2016seamless}.  In fact, systems using volunteers
exclusively, exhibit a certain amount of unpredictability, and they
might be better used in combination with ready-available computing
power.

\section{Gamification}
\label{sec:gamification}
A definition given by Huotari \cite{huotari2012defining} is ``Gamification is
the process by which gaming concepts are brought to real world tasks associated with
real people''. Gamification uses game design elements out of the domain of games
with the objective of enhancing the user\'s experience, engagement, productivity,
learning, among others. Deterding et al. proposes the following definition:
 ``Gamification” is the use of game design elements in non-game contexts''
 \cite{deterding2011gamification}.

Gamification techniques in a volunteer context seeks to persuade
users to use their natural desire to compete, learn and socialize in
given non-game context application \cite{deterding2011game,hamari2014does}.
% Some works give a form of reward to users, these include
% points \cite{sutter2010browse}, achievement badges or levels \cite{hamari2011framework},
% the filling of a progress bar \cite{o2010get}, or providing the user with virtual currency.
By Making the rewards for  tasks achievements visible to other players or
providing leader boards are ways of encouraging players to compete
\cite{hickman2010total}. In this work, a leader board was implemented
in order to promote competition, presenting only the all-time top five
participating users.
% Competition can also have problematic consequences, which can result in
% negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
% such as women \cite{kumar2013gamification}.

% \begin{figure*}[htb]
%     \centering
%         \includegraphics[width=5in]{img/login.png}
%     \caption{ User interface: Landing page, and Log-in and registration dialogs.
%     }
%     \label{fig:login}
% \end{figure*}

\section{Experimental Setup and Execution} %Pablo: change this to Experimental Setup, as it also describes the system
% OK - Mario
\label{sec:experiments}

In this experiment a gamification technique is applied to
the {\sf NodIO}
volunteer computing framework, using the
particular version described in \cite{DBLP:conf/gecco/MereloCGCRV16,2016arXiv160101607M}.
The objective of the experiment is to test the kind of impact
applying a gamification technique has on user engagement in this
particular application. In order to apply a rewarding system
user authentication had to be developed first. In earlier designs
this functionality was not desired because it was seen as a barrier
for participation. After all, the advantage of a browser based volunteer system
is precisely the minimum amount of user intervention needed to start.
The last thing a user wants to see is yet another registration form.
Then, the first design decision for this version is that registration
is optional and simple.
%as shown in Figure \ref{fig:login}.

The gamification technique employed in this work is based on a rewarding mechanism
\cite{dubois2013understanding}. In general rewards  consist of a reputation system
with score points, levels and leader boards. Points are awarded to users in response of
the accomplishment of certain activities that need to be encouraged. In these case
a point is awarded for each HTTP {\tt PUT} Request sent to the server. Levels are a long
term achievement, in this case the level depends on the score:
\[ \text{level}(\text{score};a,b)=
    \begin{cases}
      0,                                    &  \text{score}\leq a\\
      2(\frac{\text{score}-a}{b-a})^{2},    &  a\leq \text{score}\leq \frac{a+b}{2}\\
      1-2(\frac{\text{score}-b}{b-a})^{2},  & \frac{a+b}{2} \leq \text{score}\leq b\\
      1,                                    & \text{score}\geq b
   \end{cases}
\]
The function returns a normalized value that is multiplied by the maximum level
in this case 100. The variables $a$=100 and $b$=7360 are set to give users a rapid increase of
levels at the beginning.%; this function is represented in Figure \ref{fig:s}.
%
% \begin{figure*}[htb]
%     \centering
%         \includegraphics[width=3in]{img/s.png}
%     \caption{$S$ function used to assign the user's level.
%     }
%     \label{fig:s}
% \end{figure*}

After login to the NodIO web application, %Pablo: after login where? Add a link to the page or explain that the system was run in our servers
registered users are welcomed by their name, and they can see
their current score and level. All users can see the
leader board in a modal dialog.
While the modal is opened, the scores are refreshed every second. All other
functionality is available to all users and is the same as the
previous {\sf NodIO} version, showing the state of the current algorithm.
In the front page a link to the open source code of the experiment was also
available \url{https://github.com/lucero21/login-master}. %Pablo: I suggest to move the link to the introduction, just after the first time it is mentioned as free software

\subsection{Problem}
In the browser, each page visiting the experiment loads an HTTP Web Worker
that runs a local island of an evolutionary algorithm to solve a
multi-modal problem called {\em l-trap} which has been used extensively
as a benchmark for evolutionary algorithms
% and is represented in figure \ref{fig:trap},
\cite{nijssen2003analysis}.
This function counts the number of bits in a sequence $l$ and assigns
the local maximum $a$ if it has 0 bits and the global maximum $b$ if it has $l$
bits, this makes the fitness fall into a {\em trap}
as the number of bits increases, and decreasing linearly until a change in slope
is reached at point $z$, adding a deceptive component for evolutionary
algorithms. In order to increase difficulty, trap functions can be
concatenated, in our case we have used $40$ concatenated traps. The trap
function is defined as:
\[ f(u)=
    \begin{cases}
      \frac{a}{z}(z-u) & \text{if } u\leq z\\
      \frac{b}{l-z} (u-z)& \text{otherwise}
   \end{cases}
 \]
 %
% \begin{figure*}[htb]
%     \centering
%         \includegraphics[width=3in]{img/trap.png}
%     \caption{Trap function.
%     }
%     \label{fig:trap}
% \end{figure*}

Each local GA had the following parameters, the initial population was randomly
generated with a size between 128 and 256 individuals, the period to send individuals to the server was set at 100 local generations, the parameters for the trap function are fixed at $l$ = 4,
$a$=1, $b$ = 2, $z$=3 with a chromosome length of 160 bits.

%Pablo: this should be a new subsection, as we are not describing the problem from here
% Ok -Mario
\subsection{Experiment Execution}
We made a call to participation on November 23th,
2016 through the authors' social networks: ``Asking again for your help, we are conducting a computational experiment
that requires computer power. Can we borrow some of your CPU? just visit the web page (link)
and leave the tab open. Be part of the TOP TEN, register so we can track your participation.
The experiment will run until November 27th, Thanks!.'' As you can
see, the message mentioned a deadline; a link that could be used to
register was also included. Users visiting the page after the deadline
were presented with a thank you message and a final leader
board. Considering the number of friends, followers and the organic
sharing of the posts, we estimate the post was visited by around 2000
users. Only some of them registered, but in fact relatively small samples of users ($10\pm$) are sufficient for
discovering 80\% of usability problems \cite{Schmettow2012}. In fact,
we were not so much interested in making a precise model but on how
the behavior of the registered users was different from non-registered
ones; in non-parametric comparisons, using 15 sample units is in general
good enough, as seems to be the case in this particular
experiment, where we were also interested in computing the number of
users who actually registered, proving that, in fact, for most users
the fact that they have to register represents a bigger effort they
seem to be able to muster for this kind of experiments.

\subsection{Results}
\label{sec:results}

% \begin{table}[h!tbp]
%   \small
%   \caption{ Log file record structure}
%   \label{tab:record}
%   \centering
%   \small
%   \begin{tabular}{l  l}
%     \hline\noalign{\smallskip}
%     Attribute & Value \\
%     \noalign{\smallskip}\hline\noalign{\smallskip}
%     chromosome   & The individual sent to the server as a string of 160 bits.  \\ \hline
%     fitness & fitness as an integer.  \\ \hline
%     IP & IP address of the participant (later anonymized) .\\ \hline
%     user & user-name or the string `anonymous'.  \\ \hline
%     worker\_uuid & unique identifier of the HTTP Worker that sent the participation.   \\ \hline
%     level &  level of user or the string `info' \\ \hline
%     message & Only PUT messages were recorded in this experiment. \\ \hline
%     timestamp & A time-stamp based on the Unix epoch\\ \hline
%   \end{tabular}
% \end{table}
%
At the end of the experiment the resulting log file
%(whose fields are shown in Table \ref{tab:record})
contained 933,513 contributions considering both
registered and anonymous volunteers. Registered users were responsible for about
90\% of the contributions, while being only 16\% of the total number of unique users (assuming each IP is from a single user).
% The data recorded in JSON format for every contribution is  presented
% in Table~\ref{tab:record}.

\begin{figure}[h!tb]
    \centering
        \includegraphics[width=0.9\linewidth]{img/puts.png}

    \caption{Usage by registered (green triangles) and anonymous (red dots) users.  Users are ranked by
        the number of requests sent, in the \emph{x} axis is the rank, and in the \emph{y} axis
        number of PUTs in a logarithmic scale.}
      \label{fig:puts}
% Maybe we should scale dividing by number of users and put them
% together. - JJ
\end{figure}
%
Contributions are counted by the number of HTTP {\tt PUT} requests
sent to the server. %Pablo: the PUT request concept was mentioned before the level scores description. This explanation should be moved there.
In Figure~\ref{fig:puts} a comparison between the
contribution of registered vs anonymous users is presented.

The first observation arises from the length of the $x$ axis. There
were only 18 registered users vs 91 distinct IPs; the
exact number of anonymous users is upper-bounded by that amount but is
not known precisely as we only recorded the IP of the request, we did
not use cookies or other means of identification since it was not
actually needed for the experiment. Besides, a registered user
could sometimes be anonymous too, which could be caused by the fact that the first time they visited
the application they automatically begin to work as anonymous, so the
number of non-anonymous users, 91, is actually the amount of unique
IPs that had `anonymous' as the user. In the plot of
anonymous users, some of the lasts in the rank could in fact be registered users,
momentarily participating as anonymous. Even if this is not considered, only
about 16\% of users attending the call decided to participate as named
users and thus enter the game dynamics.

The second observation is apparent in the $y$ axis of Figure
\ref{fig:puts}. There is a notable difference in the amount of
participation between the two groups, and also in the slope of the
plot.  An independent-samples t-test was conducted to compare
participation, and there was a significant difference in the scores
for anonymous users  (M=1057.97, SD= 3669.86) and non-anonymous users
(M=46513.55, SD=93537.20); t-test= 4.70, p = 7.5e-06; that is, the
contribution of registered users was two orders of magnitude bigger
than for anonymous users. This difference is highlighted in
%
%Maybe a statistical comparison? Median, for instance - JJ
%count          mean           std  min      25%      50%      75%
%category
%IP        91.0   1057.978022   3669.862398  1.0     1.50     18.0    173.0
%user      18.0  46513.555556  93537.207948  7.0  3962.75  13034.5  46993.5
%
%
%      max
%category
%IP         26593.0
%user      396081.0
%Ttest_indResult(statistic=4.7070787365897289, pvalue=7.565505602664807e-06)
%
%
\begin{figure*}[htb]
    \centering
        \includegraphics[width=5in]{img/puts_box.png}
    \caption{Box-plot of the number of requests sent by registered users and anonymous IPs.
     In the \emph{y} axis are number or PUTs in a logarithmic scale.
    }
    \label{fig:box}
\end{figure*}
%
 Figure \ref{fig:box}, which is a boxplot of the  participation by two
 groups, shows the big difference between registered users (shown as
 ``users'') and anonymous ones (shown as ``IPs''). This chart also
 shows the number of outliers in the groups, which is an obvious
 consequence of the ephemeral and spontaneous nature of the
 collaboration; in fact, two registered
outliers were very important participants, maybe competing between
them, an expected consequence of the gamification, in fact, although
not in this precise form and quantity. But
there were more outliers between anonymous IPs, and some of them could
be related to the same users. The lower quartile in the users box-plot
has more
spread, indicating a participation similar to the median of anonymous
IPs, indicating that the difference in distribution is mainly due to a
few highly engaged and registered users.

%
\begin{figure*}[htb]
    \centering
        \includegraphics[width=5.3in]{img/workers_put_ip.png}
    \caption{ Participation of registered users. Each circle represents a user relating
        in the \emph{x} axis the number of workers used and in the \emph{y} axis
        number or PUTs in a logarithmic scale. The area of the circle is the number
        of unique IPs associated to every user.
    }
    \label{fig:worker-put-ips}
\end{figure*}
%
Figure~\ref{fig:worker-put-ips} gives an interesting view on the amount of
resources shared by registered users. The area of the circle is proportional
to the number of unique IPs used by each user; the user with the most participation
used a total of 66 different IPs. this number could be related with the number
of different devices used during the participation. There are some users that using less
devices participated more, this could mean a more powerful device, faster
Internet connection or simply more time spent participating. At any
rate, the chart seems to indicate that the main mechanism registered
users employ to increase their rank seems to be using all devices
available, and maxing out their CPUs by increasing the number of
workers. This indicates that, in fact, registered users who have
incurred in additional time expenses by performing registration and
login are ready to work even more by manually deploying more workers
and using more computers, which is another proof of the
socio-technical nature of volunteer computing systems.

\begin{figure*}[htb]
    \centering
        \includegraphics[width=5.3in]{img/puts_by_time.png}
    \caption{Number of PUTs by User (green dots) or IP (blue triangles) in 15 minute slots.
    }
    \label{fig:puts-time}
\end{figure*}
%
User engagement is related to the amount of time a user spends in the
application. Figure~\ref{fig:puts-time} shows two important aspects of
engagement, the overall time spent and the amount of resources shared
in that time. Registered users were more engaged during the experiment.
They contribute more resources during longer periods of time.

\begin{figure*}[htb]
    \centering
        \includegraphics[width=5in]{img/workers_best_user.png}
    \caption{ Number of workers used by the top ranked user in 15 minute slots.
    }
    \label{fig:top-user}
\end{figure*}
%
The amount of participation of the top ranked user is presented in 15 seconds time slots in
Figure~\ref{fig:top-user}. In the beginning this user was using more than 30
workers, this means that more than 30 tabs of the page were open at the
same time, perhaps using several computing devices. Some users reported that
they wanted to test the limits of their own systems, checking the percentage of
CPU they were using.

\section{Conclusions, discussion and future work}

In this paper we have added gamification to a volunteer computing
system via a leaderboard of contributions of registered users. The objective of this work was to test the kind of impact
applying a gamification technique has on user engagement in
the {\sf NodIO}
volunteer evolutionary computing framework; in concordance with
the results obtained in other
browser-based volunteer systems, after
applying the gamification techniques, registered users participated much more intensely than anonymous ones.

In fact, we have proved that, although only a minority of the users opt
for registration, their individual contributions are on average much
higher than for non-authenticated/non-gamified users. The fact that less
than one fifth of users actually registering would discard the
exclusive use of this feature; however, it is an interesting addition
to the volunteer system that in fact enhances its computing power.

The fact that applying a gamification techinque improved user participation highlights the social nature of volunteer computing systems, piling on the fact that successful {\em social clouds} must embed games and other social mechanisms to enhance participation. That is why future lines of work will look for intermediate ways of including
gamification without registration, as well as some ways of making the
user increase control over the algorithm that would go beyond simply
staying for longer or leaving the page. Since it is a sociotechnical
system, including social features in it and studying emerging social
networks is an interesting line of work too.

Other factors, such as the language used, the way the experiment is
announced, are also very important in this social context, which is
why they will be included in future lines of work too, along with the
study of the behavior of users as they are participating
in the web system.

Another line of work would be to study the possible negative effects of using
gamification techniques to improve engagement, like cheating or
literally {\em gaming} the system to defeat competition. We already
found some hints of this behavior, but more subtle effect could be taking place.
Finally, the refinement of the proposed framework will need
more case studies and further multi-disciplinary research.

\section*{Acknowledgments}

This work has been supported in part by: Ministerio espa\~{n}ol de
 Econom\'{\i}a y Competitividad under projects TIN2014-56494-C4-3-P
 (UGR-EPHEMECH) and TIN2017-85727-C4-2-P (UGR-DeepBio).

\bibliographystyle{splncs03}

\bibliography{../../bib/biblio,../../bib/evospace-i,../../bib/volunteer,../../bib/geneura}

\end{document}